---
title: "PiP-P2P Mediation Analysis"
author:
  - name: "Joshua F. Wiley"
    email: "joshua.wiley@monash.edu"
    affiliation: "Monash University"
date: "2026-03-24"
toc: true
number-sections: true
format: 
  html: 
    code-fold: true
    self-contained: true
---

# SETUP

## PACKAGES

Load required packages and a few setup functions.

```{r}
#| label: packages
#| warning: false
#| message: false

## data and modeling
library(data.table)
library(splines)
library(brms)
library(mgcv)
library(mediation)
library(mice)
library(posterior)

## for formating and rendering
library(knitr)
library(quarto)

```

Some options.

```{r}
#| label: setup

## set knitr to never cache a chunk
## this is to ensure that the code is always run and not cached
knitr::opts_chunk$set(cache = FALSE)

options(stringsAsFactors = FALSE)

## set the timezone to Melbourne
Sys.setenv(TZ = "Australia/Melbourne")

```


```{r}
#| label: simulation functions

zscore <- function(x) as.numeric(scale(x))

# Monotone piecewise-linear spline:
# positive coefficients => nonlinear and always increasing.
positive_spline <- function(x,
                            knots = c(-0.5, 0.5),
                            coefs = c(0.30, 0.20, 0.15)) {
  x <- as.numeric(x)
  coefs[1] * x +
    coefs[2] * pmax(x - knots[1], 0) +
    coefs[3] * pmax(x - knots[2], 0)
}

mediator_score <- function(m, m_type, trials = NULL) {
  switch(
    m_type,
    continuous = zscore(m),
    bernoulli  = zscore(m),
    binomial   = zscore(m / trials),   # use proportion because trials vary
    count      = zscore(log1p(m))
  )
}

mediator_trt_effect <- function(trt, m, m_type, trials = NULL) {
  if (m_type == "binomial") {
    mean(m[trt == 1] / trials[trt == 1]) - mean(m[trt == 0] / trials[trt == 0])
  } else {
    mean(m[trt == 1]) - mean(m[trt == 0])
  }
}

# ---------- mediator model ----------

make_mediator <- function(DT, m_type) {
  n   <- nrow(DT)
  trt <- DT$treatment
  s1  <- DT$strata_1_d1
  s2  <- DT$strata_2_d1
  y0  <- DT$baseline_y

  if (m_type == "continuous") {
    mu <- -0.10 +
      0.65 * trt +
      0.20 * s1 + 0.20 * s2 +
      0.35 * y0 +
      positive_spline(y0, knots = c(-0.5, 0.5), coefs = c(0.25, 0.15, 0.10))

    return(list(
      m = mu + stats::rnorm(n, mean = 0, sd = 1.0),
      trials = NULL
    ))
  }

  if (m_type == "bernoulli") {
    lp <- -0.25 +
      0.80 * trt +
      0.20 * s1 + 0.20 * s2 +
      0.55 * y0

    p <- stats::plogis(lp)

    return(list(
      m = stats::rbinom(n, size = 1, prob = p),
      trials = NULL
    ))
  }

  if (m_type == "binomial") {
    trials <- sample(6:9, size = n, replace = TRUE)

    lp <- -0.15 +
      0.70 * trt +
      0.20 * s1 + 0.20 * s2 +
      0.50 * y0

    p <- stats::plogis(lp)

    return(list(
      m = stats::rbinom(n, size = trials, prob = p),
      trials = trials
    ))
  }

  if (m_type == "count") {
    eta <- 0.00 +                              # log(1.0)
      0.35 * trt +
      0.15 * s1 + 0.15 * s2 +
      0.22 * y0 +
      positive_spline(y0, knots = c(-0.5, 0.5), coefs = c(0.18, 0.12, 0.08))

    lambda <- exp(pmin(eta, 2.0))

    return(list(
      m = stats::rpois(n, lambda = lambda),
      trials = NULL
    ))
  }

  stop("Unknown mediator type: ", m_type)
}

# ---------- outcome model ----------

make_outcome_generator <- function(DT, m, m_type, y_type, trials = NULL) {
  n   <- nrow(DT)
  trt <- DT$treatment
  s1  <- DT$strata_1_d1
  s2  <- DT$strata_2_d1
  y0  <- DT$baseline_y

  mz <- mediator_score(m, m_type, trials)

  # Nonlinear mediator -> outcome association only when BOTH mediator and outcome
  # are continuous/count.
  nonlinear_pair <- m_type %in% c("continuous", "count") &&
    y_type %in% c("continuous", "count")

  if (nonlinear_pair) {
    m_term <- 0.80 * zscore(
      positive_spline(
        mz,
        knots = c(-0.25, 0.75),
        coefs = c(0.30, 0.20, 0.15)
      )
    )
  } else if (y_type == "continuous") {
    m_term <- 0.60 * mz
  } else if (y_type == "bernoulli") {
    m_term <- 0.85 * mz
  } else if (y_type == "count") {
    m_term <- 0.45 * mz
  } else {
    stop("Unknown outcome type: ", y_type)
  }

  # Fixed random draws so calibration of cor(baseline, outcome) is stable.
  eps_cont <- stats::rnorm(n, mean = 0, sd = 1.0)
  u_bin    <- stats::runif(n)
  u_cnt    <- stats::runif(n)

  function(beta_y0) {
    if (y_type == "continuous") {
      eta <- 0.10 * trt +
        0.12 * s1 + 0.12 * s2 +
        m_term +
        beta_y0 * y0

      return(eta + eps_cont)
    }

    if (y_type == "bernoulli") {
      eta <- -0.20 +
        0.10 * trt +
        0.12 * s1 + 0.12 * s2 +
        m_term +
        beta_y0 * y0

      p <- stats::plogis(eta)
      return(as.integer(u_bin < p))
    }

    if (y_type == "count") {
      eta <- 0.15 +
        0.08 * trt +
        0.10 * s1 + 0.10 * s2 +
        m_term +
        beta_y0 * y0

      lambda <- exp(pmin(eta, 2.0))
      return(stats::qpois(u_cnt, lambda = lambda))
    }
  }
}

# Choose beta for baseline_y so cor(baseline_y, post outcome) is ~ 0.5.
choose_beta_y0 <- function(gen_y,
                           y0,
                           target_r = 0.50,
                           grid = seq(0, 2.5, by = 0.05)) {

  cors <- vapply(
    grid,
    FUN.VALUE = numeric(1),
    FUN = function(b) {
      y <- gen_y(b)
      if (stats::sd(y) == 0) return(NA_real_)
      stats::cor(y0, y)
    }
  )

  ok <- which(!is.na(cors))
  best <- ok[which.min(abs(cors[ok] - target_r))]

  list(
    beta_y0    = grid[best],
    achieved_r = cors[best]
  )
}

# ---------- main simulator ----------

simulate_mediation_data <- function(n = 1000,
                                    seed = 123,
                                    target_r = 0.50) {
  set.seed(seed)

  DT <- data.table(id = seq_len(n))

  # Two fully crossed 2-level strata, dummy coded as 0/1.
  DT[, `:=`(
    strata_1 = stats::rbinom(.N, size = 1, prob = 0.5),
    strata_2 = stats::rbinom(.N, size = 1, prob = 0.5)
  )]

  DT[, `:=`(
    strata_1_d1 = as.integer(strata_1 == 1L),
    strata_2_d1 = as.integer(strata_2 == 1L)
  )]

  DT[, stratum := factor(paste0("S", strata_1, strata_2))]

  # Single-level treatment assignment, balanced within strata cells.
  DT[, treatment := sample(rep(c(0L, 1L), length.out = .N)),
     by = .(strata_1, strata_2)]

  # Baseline outcome covariate.
  DT[, baseline_y := zscore(
    stats::rnorm(.N) +
      0.25 * strata_1_d1 +
      0.25 * strata_2_d1 +
      stats::rnorm(.N, sd = 0.75)
  )]

  # Coding:
  # mediator: 1=continuous, 2=bernoulli, 3=binomial, 4=count
  # outcome : 1=continuous, 2=bernoulli, 3=count
  mediator_types <- c("continuous", "bernoulli", "binomial", "count")
  outcome_types  <- c("continuous", "bernoulli", "count")

  design <- CJ(m_idx = seq_along(mediator_types),
               y_idx = seq_along(outcome_types))

  design[, `:=`(
    mediator_type = mediator_types[m_idx],
    outcome_type  = outcome_types[y_idx],
    mediator_name = sprintf("m_%d_%d", m_idx, y_idx),
    outcome_name  = sprintf("y_%d_%d", m_idx, y_idx),
    trials_name   = fifelse(
      mediator_types[m_idx] == "binomial",
      sprintf("m_trials_%d_%d", m_idx, y_idx),
      NA_character_
    ),
    nonlinear_pair = mediator_types[m_idx] %in% c("continuous", "count") &
      outcome_types[y_idx] %in% c("continuous", "count")
  )]

  for (k in seq_len(nrow(design))) {
    mt <- design$mediator_type[k]
    yt <- design$outcome_type[k]
    mn <- design$mediator_name[k]
    yn <- design$outcome_name[k]
    tn <- design$trials_name[k]

    med <- make_mediator(DT, m_type = mt)
    DT[, (mn) := med$m]

    if (!is.na(tn)) {
      DT[, (tn) := med$trials]
    }

    gen_y <- make_outcome_generator(
      DT     = DT,
      m      = med$m,
      m_type = mt,
      y_type = yt,
      trials = med$trials
    )

    tuned <- choose_beta_y0(
      gen_y    = gen_y,
      y0       = DT$baseline_y,
      target_r = target_r
    )

    DT[, (yn) := gen_y(tuned$beta_y0)]


    design[k, `:=`(
      beta_y0        = tuned$beta_y0,
      achieved_r     = tuned$achieved_r,
      trt_effect_on_m = mediator_trt_effect(
        trt    = DT$treatment,
        m      = med$m,
        m_type = mt,
        trials = med$trials
      ),
      mediator_mean  = if (mt == "binomial") mean(med$m / med$trials) else mean(med$m),
      outcome_mean   = mean(DT[[yn]])
    )]
  }

  list(
    data   = DT,
    design = design
  )
}

```

```{r}
#| label: mediation functions

med_summary <- function(obj) {
    rbind(
        c(Estimate = obj[["d0"]], obj[["d0.ci"]]),
        c(Estimate = obj[["d1"]], obj[["d1.ci"]]),
        c(Estimate = obj[["d.avg"]], obj[["d.avg.ci"]])
    )
}

nie_brms <- function(fit,
                     treatment_var = "A",
                     mediator_response = "M",
                     mediator_predictor = "M",
                     outcome_response = "Y",
                     scaling_variable = NULL,
                     k = 1000L) {
    tmp <- as.data.table(model.frame(fit))

    if (!treatment_var %in% names(tmp)) {
        stop("Treatment variable not found in model frame: ", treatment_var)
    }
    if (!mediator_response %in% names(tmp)) {
        stop("Mediator variable not found in model frame: ", mediator_response)
    }
    if (!mediator_predictor %in% names(tmp)) {
        stop("Mediator variable not found in model frame: ", mediator_predictor)
    }
    if (!outcome_response %in% names(tmp)) {
        stop("Outcome variable not found in model frame: ", outcome_response)
    }
    if (!is.null(scaling_variable) ) {
          if (!scaling_variable %in% names(tmp)) {
              stop("Scaling variable not found in model frame: ", scaling_variable)
          }
    }

    txindex <- rep(c(0L, 1L), each = nrow(tmp))
    tmp <- rbind(tmp, tmp)
    tmp[txindex == 0, (treatment_var) := 0]
    tmp[txindex == 1, (treatment_var) := 1]

    tmp2 <- rbind(tmp, tmp)
    tmp2[, (mediator_response) := as.numeric(get(mediator_response))]
    # typeof(tmp2[[mediator_response]])
    txindex2 <- c(txindex, txindex)

    nk <- ndraws(fit)
    if (nk >= k) {
        index <- round(seq(from = 1, to = nk, length.out = k))
        if (!identical(anyDuplicated(index), 0L)) {
            warning("There were duplicated draws.")
        }
    } else {
        stop("Number of posterior draws (", nk, ") is less than k (", k, ").")
    }

    # Posterior mediator predictions
    mdraw <- posterior_predict(
        fit,
        newdata = tmp, draw_ids = index,
        resp =  gsub("\\.|_", "", make.names(mediator_response, unique = TRUE))
    )

    results <- data.table(
        NIE0 = rep(NA_real_, length(index)),
        NIE1 = rep(NA_real_, length(index)),
        NIE_avg = rep(NA_real_, length(index))
    )

    for (i in seq_along(index)) {
        ui <- index[i]

        if (is.null(scaling_variable)) {
            tmpm <- mdraw[i, ]
        } else {
            tmpm <- mdraw[i, ] / tmp[[scaling_variable]]
        }
        
        tmp2[txindex2 == 0, (mediator_predictor) := tmpm]
        tmp2[txindex2 == 1, (mediator_predictor) := tmpm]

        yhat <- posterior_predict(
            fit,
            newdata = tmp2,
            draw_ids = ui,
            resp =  gsub("\\.|_", "", make.names(outcome_response, unique = TRUE))
        )

        results[i, NIE0 := mean(
            yhat[txindex2 == 0][txindex == 1] -
                yhat[txindex2 == 0][txindex == 0]
        )]
        results[i, NIE1 := mean(
            yhat[txindex2 == 1][txindex == 1] -
                yhat[txindex2 == 1][txindex == 0]
        )]

        results[i, NIE_avg := (NIE0 + NIE1) / 2]
    }
    return(results)
}

```

```{r}
#| label: run-simulation
#| results: asis

# Examples:
# m_1_1 = continuous mediator, continuous outcome y_1_1
# m_1_2 = continuous mediator, bernoulli outcome y_1_2
# m_4_3 = count mediator, count outcome y_4_3
#
# For binomial mediators, the matching trial count column is:
# m_trials_3_1, m_trials_3_2, m_trials_3_3

out <- simulate_mediation_data(n = 2000, seed = 2026, target_r = 0.50)

kable(out$design, 
  digits = 2,
  caption = "Various mediator and outcome types in simulated data.")

```

This is an example (not run) that can be used to compare 
the `brms`-based NIE estimation with the `mediate` package.

```{r}
#| label: comparison mediation
#| eval: false

brm_2_3 <- brm(
    bf(m_2_3 ~ treatment + strata_1_d1 + strata_2_d1 + baseline_y) +
        bf(y_2_3 ~ treatment + m_2_3 + strata_1_d1 + strata_2_d1 + baseline_y) +
        set_rescor(FALSE),
    data = out$data,
    family = list(bernoulli(), poisson()),
    chains = 4, iter = 4000, warmup = 2000,
    cores = 4, seed = 20260324L, backend = "cmdstanr",
    control = list(adapt_delta = 0.95)
)

summary(brm_2_3)

indeff_2_3 <- nie_brms(
    fit = brm_2_3,
    treatment_var = "treatment",
    mediator_response = "m_2_3",
    mediator_predictor = "m_2_3",
    outcome_response = "y_2_3",
    k = 1000L
)

posterior_summary(indeff_2_3)

## Design 2_3 with classical GLMs for the `mediate` package.
glm_m_2_3 <- glm(
    m_2_3 ~ treatment + strata_1_d1 + strata_2_d1 + baseline_y,
    data = out$data,
    family = binomial(link = "logit")
)

glm_y_2_3 <- glm(
    y_2_3 ~ treatment + m_2_3 + strata_1_d1 + strata_2_d1 + baseline_y,
    data = out$data,
    family = poisson(link = "log")
)

med_2_3 <- mediate(
    model.m = glm_m_2_3,
    model.y = glm_y_2_3,
    treat = "treatment",
    mediator = "m_2_3",
    sims = 1000
)

summary(med_2_3)
```

We expect the mediator to be a binomial variable and the trial outcomes
are typically continuous and expected to be approximately normally distributed. So the most relevant design is 3_1, which is what we focus on here.

```{r}
#| label: mediation-fit

# Design 3_1
out$data$prop_m_3_1 <- out$data$m_3_1 / out$data$m_trials_3_1

brm_3_1 <- brm(
    bf(m_3_1 | trials(m_trials_3_1) ~ treatment + strata_1_d1 + strata_2_d1 + baseline_y) +
        bf(y_3_1 ~ treatment + s(prop_m_3_1, k = 3) + strata_1_d1 + strata_2_d1 + baseline_y) +
        set_rescor(FALSE),
    data = out$data,
    family = list(binomial(), gaussian()),
    chains = 4, iter = 2000, warmup = 1000,
    cores = 4, seed = 20260324L, backend = "cmdstanr",
    control = list(adapt_delta = 0.95),
    refresh = 0
)

summary(brm_3_1)

indeff_3_1 <- nie_brms(
    fit = brm_3_1,
    treatment_var = "treatment",
    mediator_response = "m_3_1",
    mediator_predictor = "prop_m_3_1",
    outcome_response = "y_3_1",
    scaling_variable = "m_trials_3_1",
    ## just 100 for speed but should be higher
    ## but should be higher for stable / precise estimates
    k = 100L 
)

```

```{r}
#| label: mediation-output
#| results: asis

kable(posterior_summary(indeff_3_1), 
  digits = 2,
  caption = "NIE (ACME) for control (NIE0), treatment (NIE1), and average (NIE_avg) with 95% credible intervals.")

```


# Design 3_1 With Missing Data and Multiple Imputation

To illustrate a simple missing-data workflow for the main `3_1` design, the code
below first imposes about 10% missingness on the outcome, then uses
predictive mean matching via `mice` to create 10 imputed datasets using only the
variables from the analytic model. Each imputed dataset is then analysed with the
same joint `brms` model, `nie_brms()` is run with `k = 100L` per dataset, and
the posterior draws are pooled across imputations.
This code is not evaluated as it is time consuming to run.

```{r}
#| label: mediation-missing-data
#| eval: false

## Start from only the variables used in the analytic model.
mi_data_3_1 <- copy(out$data[, .(
    treatment,
    strata_1_d1,
    strata_2_d1,
    baseline_y,
    m_3_1,
    m_trials_3_1,
    y_3_1
)])

## Add about 10% missingness to the outcome.
set.seed(20260324L)
mi_data_3_1[sample(.N, size = ceiling(.N * 0.10)), y_3_1 := NA_real_]

## Set up mice so only variables in the analytic model are used.
mi_method <- mice::make.method(mi_data_3_1)
mi_method[] <- ""
mi_method["y_3_1"] <- "pmm"

mi_predictor_matrix <- mice::make.predictorMatrix(mi_data_3_1)
mi_predictor_matrix[,] <- 0L
mi_predictor_matrix["y_3_1", c(
    "treatment",
    "strata_1_d1",
    "strata_2_d1",
    "baseline_y",
    "m_3_1",
    "m_trials_3_1"
)] <- 1L

imp_3_1 <- mice(
    data = as.data.frame(mi_data_3_1),
    m = 10,
    method = mi_method,
    predictorMatrix = mi_predictor_matrix,
    maxit = 20,
    seed = 20260324L,
    printFlag = FALSE
)

imp_data_3_1 <- lapply(seq_len(imp_3_1$m), function(i) {
    DT <- as.data.table(mice::complete(imp_3_1, action = i))
    DT[, prop_m_3_1 := m_3_1 / m_trials_3_1]
    DT
})

fit_3_1_mi <- vector("list", length = length(imp_data_3_1))
nie_3_1_mi <- vector("list", length = length(imp_data_3_1))

for (i in seq_along(imp_data_3_1)) {
    fit_3_1_mi[[i]] <- brm(
        bf(m_3_1 | trials(m_trials_3_1) ~ treatment + strata_1_d1 + strata_2_d1 + baseline_y) +
            bf(y_3_1 ~ treatment + s(prop_m_3_1, k = 3) + strata_1_d1 + strata_2_d1 + baseline_y) +
            set_rescor(FALSE),
        data = imp_data_3_1[[i]],
        family = list(binomial(), gaussian()),
        chains = 4, iter = 2000, warmup = 1000,
        cores = 4, seed = 20260324L + i, backend = "cmdstanr",
        control = list(adapt_delta = 0.95)
    )

    nie_3_1_mi[[i]] <- copy(nie_brms(
        fit = fit_3_1_mi[[i]],
        treatment_var = "treatment",
        mediator_response = "m_3_1",
        mediator_predictor = "prop_m_3_1",
        outcome_response = "y_3_1",
        scaling_variable = "m_trials_3_1",
        k = 100L
    ))

    nie_3_1_mi[[i]][, imputation := i]
}

pooled_nie_3_1 <- rbindlist(nie_3_1_mi, use.names = TRUE)

posterior_summary(pooled_nie_3_1[, .(NIE0, NIE1, NIE_avg)])

```
